---
title: "PRSPGxTL Vignette"
author: "Youshu Cheng, Song Zhai, Wujuan Zhong, Rachel Marceau West, Judong Shen"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette
  #pdf_document: default
vignette: >
  %\VignetteIndexEntry{PRSPGxTL Vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

A transfer learning (TL) based method to leverage large-scale disease GWAS summary statistics and individual-level pharmacogenomics (PGx) data to predict drug response

We designed a total of 6 models (M1-M6) of PRS-PGx-TL depending on the following criteria:

**Updating Parameters**: 
   - Whether to update $\beta_G$ and $\beta_{GT}$ at the same time, or fix $\beta_G = \beta_G^{pre}$ and only update $\beta_{GT}$.

**Initial Values for $\beta_{GT}$**: 
   - Whether to set the initial values for $\beta_{GT}$ as 0 or $\beta_G^{pre}$.

**Criterion for Parameter Tuning**: 
   - Whether to maximize the overall $R^2$ or the conditional $R^2$ of $G \times T$ (conditional on $G$) as a criterion for parameter tuning.

The choice of M1-M6 corresponds to the following arguments:

| Model | fixG | initial values (`initial`) | criterion (`best_criterion`)
| :--- | :--- | :--- | :--- | 
| M1 | `FALSE` | `zero` | `best_R2`
| M2 | `FALSE` | `PRS` | `best_R2`
| M3 | `FALSE` | `zero` | `best_condR2`
| M4 | `FALSE` | `PRS` | `best_condR2`
| M5 | `TRUE` | `zero` | `best_R2`
| M6 | `TRUE` | `PRS` | `best_R2`

We will give an example of how to use the PRSPGxTL package. 

## Example

```{r}
library(PRSPGxTL)
```

For example, if we want to use model M2, we can use the following code.

```{r, results='hide', message=FALSE, warning=FALSE}
# Load the example data
data(sum_stats)
data(ped)
data(G)
data(bim)

# set parameters
fixG = FALSE
initial = "PRS"
best_criterion = "best_R2"

# use model M2
s1_results = inner_layer_CV(sum_stats = sum_stats, ped = ped, G = G, bim = bim, initial = initial, num_snp = 2349, covar = NULL, fixG = fixG, verbose = TRUE)
s2_results = parameter_tuning(X = ped, s1_results = s1_results, fixG = fixG)
s3_results = rerun_algorithm(sum_stats = sum_stats, ped = ped, G = G, bim = bim, initial = initial, num_snp = 2349, s2_results = s2_results, fixG = fixG)

```

Estimated $\beta_G$ and $\beta_{GT}$ can be found in `s3_results[['beta']][['beta_g']]` and `s3_results[['beta']][['beta_gt']]`, respectively.

```{r}
cat("Estimated beta_G:", "\n")
est_beta_G = s3_results[['beta']][['beta_g']]
head(est_beta_G)

cat("Estimated beta_GT:", "\n")
est_beta_GT = s3_results[['beta']][['beta_gt']]
head(est_beta_GT)
```

Constructed prognostic and predictive PRS in the testing dataset can be found in `s3_results[['testScoresum']][['score_g']]` and `s3_results[['testScoresum']][['score_gt']]`, respectively.

```{r}
cat("Constructed prognostic PRS in the testing dataset:", "\n")
est_PRS_G = s3_results[['testScoresum']][['score_g']]
head(est_PRS_G)

cat("Constructed predictive PRS in the testing dataset:", "\n")
est_PRS_GT = s3_results[['testScoresum']][['score_gt']]
head(est_PRS_GT)
```


